TEMPERATURES AND METALLICITIES OF M DWARFS IN THE APOGEE SURVEY

### 2019-10-30 paper/draft.tex
referee response

A number of studies focused on improving M dwarf metallicity precisions have used systems of M dwarfs in common proper motion with an FGK star and strong, isolated lines in the spectra of the M dwarf \citep[e.g.,][]{Terrien:2012,RojasAyala:2010,Newton:2014,Neves:2014,Lindgren:2016} to developed precise empirical relations (as good as $\sim$0.07\,dex). 
However these metallicity calibrations do not take advantage of the full wavelength coverage available, nor information about the overall spectral shape often used to determine $\teff$ and spectral type. Furthermore, earlier calibrations are generally based on moderate-resolution data \citep[with some exceptions,][]{Neves:2014, Lindgren:2016} that fail to utilize the greater spectral information provided by APOGEE's resolution.

@@

\thecannon\ is a regression model which relies on two assumptions: first, that sources with identical labels have near-identical flux at each wavelength pixel; and second, that the expected flux at each pixel varies continuosly with change in label.

@@

Following the procedure of \citet{Ness:2015} and \citet{Ho:2017a} we adopt a simple linear model that assumes the flux at each pixel of the spectrum can be parameterized as a function of a label vector $\ell$ and coefficient vector \emph{$\theta$}. For each star \emph{n}, at wavelength pixel \emph{$\lambda$}, we assume that the measured flux for a star at a given pixel is the sum of the coefficient and label product, and observational noise:
\begin{equation}
	f_{n\lambda} = \theta_{\lambda}^{T} \cdot \ell_{n} + N_{\lambda} \label{eqn:cannon}
\end{equation} 
\color{red}{Here we use the noise model}\color{black} $N_{\lambda}=[s_{\lambda}^2 + \sigma_{n\lambda}^2]\xi_{n\lambda}$, where the bracketed term is the root mean sum of the intrinsic scatter of the model at each pixel \emph{$s_{\lambda}$}, and the uncertainty due to instrumental effects \emph{$\sigma_{n\lambda}$}, which is then multiplied times a Gaussian random number $\xi_{n\lambda} \sim \mathcal{N} (0,1)$. Equation (\ref{eqn:cannon}) corresponds to the single-pixel log-likelihood function:
\begin{equation}
	\ln p(f_{n\lambda}|\theta^T_{\lambda}, \ell_n, s^2_n) = 
	-\frac{1}{2}\frac{[f_{n\lambda} - \theta_{\lambda}^{T} \cdot \ell_{n}]^2}{s_{\lambda}^2 + \sigma_{n\lambda}^2} - \frac{1}{2} \ln(s_{\lambda}^2 + \sigma_{n\lambda}^2) 
	\label{eqn:likelihood}
\end{equation}
which gives the probability density function of the measured flux, given the labels, coefficients and scatters.

@@

\thecannon\ model can in principle be trained on any physical or empirical labels available, giving a wide range of flexibility to the model. However, choosing a training sample with high quality labels is critical to its performance.
Limitations of \thecannon\ include that test (output) labels are only accurate if the training labels are accurate, and only precise if the training labels are measured consistently across the training sample.
It is also critical to have a training sample with the dynamic range to span the entire parameter space of interest, as \thecannon\ does not extrapolate well outside the parameter space of the training sample.
Finally, \thecannon\ assumes that the dependence of \color{red}{the}\color{black} spectrum on labels is continuous and smooth---and in this implementation, well approximated by quadratic functions. If that isn't true, there will be features that \thecannon\ cannot reproduce.

For the purpose of this study, we have constructed two different training samples: first a one-dimensional \emph{spectral type model}, and second, a two-dimensional \emph{physical parameter model}, which describes the temperature and metallicity. The choice of training labels, dimensionality of our data set, and requirements for a good training set \color{red}{are discussed}\color{black} further in Section \ref{sec:discussion}.

@@

The physical parameter training sample consists of 87 sources with reference labels distributed over $2859 <T_{\rm eff} < 4131$K, and $-0.48 < \feh < 0.49$\,dex, 41 sources of which are drawn from \citet{Mann:2015} (hereafter M15), and 46 of which are part of a previously unpublished extension sample to M15, analyzed using similar data and identical techniques to M15. The major difference between the extension sample \color{red}{to M15}\color{black} is that the sample had lower-quality or no parallaxes (prior to Gaia data) and hence were omitted from the M15 study and were less vetted for binarity than the M15 sample.

@@

Another mode of analysis we can look at with \thecannon\ is how the derivative of the model changes with respect to given training parameters, which makes our model interpretable for discovering or verifying atomic or molecular lines with strong dependence on different physical parameters. The top two panels of Figures \ref{fig:demo_teff} and \ref{fig:demo_feh} show two example spectra and model fits for two different temperatures (Figure \ref{fig:demo_teff}), and two different metallicities (Figure \ref{fig:demo_feh}), with atomic and molecular features identified by the abundance analysis of \citet{Souto:2017}. The bottom panels of Figures \ref{fig:demo_teff} and \ref{fig:demo_feh} show the derivative of flux with respect to temperature and metallicity at each pixel, taken at the median training values. In order to evaluate which spectral features show statistically significant change with respect to input label, we compute \color{red}{the}\color{black} error of the derivative at each pixel using a jackknife statistic (with a 1-$\sigma$ level overplotted in red):
\begin{equation}
	\sigma_{\theta,m}^2 = \frac{N-1}{N} \sum^N_{n=1} (\theta_{/n}^m - \theta^m)^2 
\end{equation}
where $\sigma_{\theta,m}$ is the error at pixel \emph{m}, \emph{N} is the total number of stars in the sample indexed by \emph{n}, $\theta$ is the coeffient vector trained on all \emph{N} sources, and $\theta_{/n}$ is the coefficient vector trained on \emph{N-1} sources excluding star \emph{n}. A summary of identified lines with derivative values greater than $2\sigma_{\rm jackkife}$ is given in Table \ref{table:teff_derivative} and Table \ref{table:feh_derivative}.

@@

The spectra contain roughly 8000 pixels,
so we might expect the $\chi^2$ values to be close to 8000 in magnitude,
\color{red}{however, they}\color{black} are much higher.
This discrepancy flows from the fact that, while the spectral model is good at the percent
level, the signal-to-noise of the typical spectra are more than 100.
That is, the $\chi^2$ values do show that the model is not good in the frequentist
sense; it is only good at the percent level.

@@

We trained \thecannon\ on 51 M dwarfs ranging \color{red}{between}\color{black} M0$-$M9 with a one-dimensional spectral type label, and obtained a precision of $\pm$0.9 spectral types, also similar to the original training label uncertainty of $\pm$1 spectral type. We note however, that the training sample is distributed heavily towards earlier type sources, with a median spectral type of 3 and only one M8 and one M9 source. As seen in Figure \ref{fig:west_validation}, the model performs poorly at reproducing spectral types $>$M8, which confirms that \thecannon\ does not extrapolate well to labels outside of the training sample space. Because of this skew for late-type sources, we report our spectral type model to be precise to $\pm$0.7 spectral types for the range M0-M6. Repeating the analysis of Section \ref{subsec:mann_results}, Figure \ref{fig:west_validation} shows LOOCV test for the labels reported in Table \ref{table:west_results}, and Figure \ref{fig:west_derivative} shows the derivative of model flux with varying spectral type.

@@

\begin{enumerate}
\item \textbf{Quality of fit cut:} We apply a \cannon\ model $\chi^2$ cut of less than 100,000, chosen to remove badly fit sources (such as fast rotators), but include $\chi^2$ values close \color{red}{to}\color{black} the distribution of training sample (Figure \ref{fig:chi_dist}).

\item \textbf{Color-magnitude cuts:} Using \gaia\ and 2MASS photometries we apply the color-magnitude selections shown in Figure \ref{fig:cmd_selection} to remove sources above the main sequence (which are likely pre-main sequence, reddened K dwarfs and\color{red}{/or}\color{black} multiples), and subdwarfs below the main sequence.

@@

\begin{figure}
\begin{center}
	\includegraphics[width=\linewidth]{wide_binary_met.png}
	\caption{Metallicities of \cannon\ M dwarfs in binaries (x-axis) compared to the metallicities of their companions (y-axis). Label colors show the source and spectral type of the companion metallicities. Metallicity values for each binary pair is compiled in Table \ref{table:binaries}. The overall scatter between the 15 metallicity pairs is 0.08\,dex.}
\end{center}
\end{figure}

@@

As a check of our test sample metallicity reliability, we cross-matched our M dwarf final sample with the \citet{Elbadry:2018b} catalog of $>$50,000 high-confidence, widely separated binaries identified by \drtwo. 
In total we found 216 of the APOGEE M dwarfs to have binary pairs (46 FGK+M, 155 M+M, and 15 WD+M). Out of the 155 M+M pairs, 8 pairs contained both pairs in APOGEE. Cross-matching the list of FGK+M dwarf companions with several catalogs/surveys with measured stellar metallicities, we found an additional 7 sources with FGK metallicities from LAMOST \citep{Zhao:2012} and APOGEE (ASPCAP). The metallicity measurements for the 15 M dwarf binaries and their companions are given in Table \ref{table:binaries}, and the overall scatter is $0.08\,$dex--an improvement over the scatter of ASPCAP metallicities which is $0.15\,$dex for these 15 sources. Internal consistency of the two models (the scatter of the 8 M+M pairs both in APOGEE) are 0.06\,dex and 0.12\,dex for \cannon\ and ASPCAP respectively.

@@

While \thecannon\ achieves excellent precision at predicting labels
and reproducing spectral features, the
accuracy of labels it produces is limited by
the accuracy, relative precision, size, dynamic range and
representation of the training sample.
That is, being a supervised method, \thecannon\ is never any better in a mean (bias) sense than the input training data, although it can be better in a precision or variance sense.
The catalog we have produced is a label transfer from parameters provided
in our input data (M15) and it implicitly adopts all the biases and
issues from that input data.
It is also limited to the stellar-parameter domain of that input catalog.
That said, this work provides and external validation of the M15 stellar parameters.

@@

While a detailed analysis of atmospheric model limitations is beyond the scope of this paper, 
our results provide an avenue to compare the FGK star metallicity scale 
to the less well-understood M dwarf metallicity scale.
These results find that atmospheric metallicities are systematically metal-poor biased
compared to \cannon-based metallicities trained on sources with metallicities calibrated
to those of FGK companions. 
At the high temperature end ($\teff > 3600$\,K), the ASPCAP metallicity bias is $-0.12-0.15\,$dex with a scatter of $0.05-0.06\,$dex relative to \cannon\ metallicities, and increases to a bias of $-0.3\,$dex and scatter of $0.13\,$dex at the lower temperature end ($\teff < 3200$\,K) (Figure \ref{fig:met_comp}).
We suspect that this metal-poor bias, while not explored to a great extent in this work, is due 
to the line lists of the models: 
an effect in which the optimizer of the pipeline may be lowering the continuum level and metallicity of the fit to compensate for the missing lines or opacities.
Further analysis would need to be done to determine if the metallicity bias
is found in other model grids (besides the ATLAS/MARCS models used by the ASPCAP pipeline), 
and if the effect is present at other wavelengths.

@@

%%==========================================================================================
\acknowledgements
We would like to acknowledge Hans Walter Rix (MPIA), Derek Homeier (Universit{\"a}t Heidelberg), Wolfgang Bradner (MPIA), Anna-Christina Eilers (MPIA), Melissa Ness (Columbia), Kevin Covey (WWU), Diogo Souto (Observatário Nacional/MCTI), Keivan Stassun (Vanderbilt), \color{red}{Katia Cunha (NOAO), Anthony Brown (Leiden), Aida Behmard (Caltech) and Christopher Theissen (UCSD)}\color{black} for constructive discussions in the process of this project, as well as Bertrand Goldman (MPIA) and those who have supported the internship program at the Max Planck Institute f{\"u}r Astronomie for providing JB with funding and hospitality.
This project was partially supported by
the US National Aeronautics and Space Administration (NASA grant NNX12AI50G),
the US National Science Foundation (NSF grant AST-1517237),
and the Moore--Sloan Data Science Environment at NYU.
A.J.B. acknowledges funding support from the National Science Foundation under award No. AST-1517177.
This work is supported by the SDSS Faculty and Student Team (FAST) initiative.

@@

\begin{figure*}
	\begin{center}
	\includegraphics[width=16cm]{demo_derivatives_teff3.png}
	\end{center}
	\caption{\textit{Top two panels of each plot:} Mann-trained model for varying temperatures, and similar metallicities. \textit{Third panel of each plot:} Derivative of \thecannon\ model with respect to temperature, taken at the median training temperature, T$_{\rm eff}=3463$K; error estimate computed using a jackknife statistic at each pixel is marked in red, making it possible to distinguish which features vary significantly with change in spectral type, and which are likely due to noise.}
	\label{fig:demo_teff}
\end{figure*}

@@

\begin{figure*}
	\begin{center}
	\includegraphics[width=16cm]{demo_derivatives_feh3.png}
	\end{center}
	\caption{\textit{Top two panels of each plot:} Mann-trained model for varying metallicities and similar temperatures. \textit{Third panel:} Derivative of \thecannon\ model with respect to metallicity, taken at the median training metallicity, $\feh=-0.03$\,dex; jackknife computed error at each pixel is shown in red.} \label{fig:demo_feh}
\end{figure*}

>>>>

A number of studies focused on improving M dwarf metallicity precisions have used systems of M dwarfs in common proper motion with an FGK star and strong, isolated lines in the spectra of the M dwarf \citep[e.g.,][]{Terrien:2012,RojasAyala:2010,Newton:2014,Neves:2014,Lindgren:2016} to develope precise empirical relations (as good as $\sim$0.07\,dex). 
However these metallicity calibrations do not take advantage of the full wavelength coverage available, nor information about the overall spectral shape often used to determine $\teff$ and spectral type. Furthermore, earlier calibrations are generally based on moderate-resolution data \citep[with some exceptions,][]{Neves:2014, Lindgren:2016} that fail to utilize the greater spectral information provided by APOGEE's resolution.

@@

\thecannon\ is a regression model which relies on two assumptions: first, that sources with identical labels have near-identical flux at each wavelength pixel; and second, that the expected flux at each pixel varies continuously with change in label.

@@

Following the procedure of \citet{Ness:2015} and \citet{Ho:2017a} we adopt a simple linear model that assumes the flux at each pixel of the spectrum can be parameterized as a function of a label vector $\ell$ and coefficient vector \emph{$\theta$}. For each star \emph{n}, at wavelength pixel \emph{$\lambda$}, we assume that the measured flux for a star at a given pixel is the sum of the coefficient and label product, and observational noise:
\begin{equation}
	f_{n\lambda} = \theta_{\lambda}^{T} \cdot \ell_{n} + N_{\lambda} \label{eqn:cannon}
\end{equation} 
\color{red}{\bf Here we use the noise model }\color{black} $N_{\lambda}=[s_{\lambda}^2 + \sigma_{n\lambda}^2]\xi_{n\lambda}$, where the bracketed term is the root mean sum of the intrinsic scatter of the model at each pixel \emph{$s_{\lambda}$}, and the uncertainty due to instrumental effects \emph{$\sigma_{n\lambda}$}, which is then multiplied times a Gaussian random number $\xi_{n\lambda} \sim \mathcal{N} (0,1)$. Equation (\ref{eqn:cannon}) corresponds to the single-pixel log-likelihood function:
\begin{equation}
	\ln p(f_{n\lambda}|\theta^T_{\lambda}, \ell_n, s^2_n) = 
	-\frac{1}{2}\frac{[f_{n\lambda} - \theta_{\lambda}^{T} \cdot \ell_{n}]^2}{s_{\lambda}^2 + \sigma_{n\lambda}^2} - \frac{1}{2} \ln(s_{\lambda}^2 + \sigma_{n\lambda}^2) 
	\label{eqn:likelihood}
\end{equation}
which gives the probability density function of the measured flux, given the labels, coefficients and scatters.

@@

\thecannon\ model can in principle be trained on any physical or empirical labels available \color{red}{\bf beyond those which typically parameterize theoretical atmospheric models ($\teff$, logg, [M/H], etc.), such as additional physical parameters (e.g., mass/age \citealt{Ness:2016}) or empirical proxies for physical parameters (e.g., spectral types, colors, magnitudes), }\color{black} giving a wide range of flexibility to the model. 
However, choosing a training sample with high quality labels is critical to its performance.
Limitations of \thecannon\ include that test (output) labels are only accurate if the training labels are accurate, and only precise if the training labels are measured consistently across the training sample.
It is also critical to have a training sample with the dynamic range to span the entire parameter space of interest, as \thecannon\ does not extrapolate well outside the parameter space of the training sample.
Finally, \thecannon\ assumes that the dependence of \color{red}{\bf the }\color{black} spectrum on labels is continuous and smooth---and in this implementation, well approximated by quadratic functions. If that isn't true, there will be features that \thecannon\ cannot reproduce.

For the purpose of this study, we have constructed two different training samples: first a one-dimensional \emph{spectral type model}, and second, a two-dimensional \emph{physical parameter model}, which describes the temperature and metallicity. The choice of training labels, dimensionality of our data set, and requirements for a good training set \color{red}{\bf are discussed }\color{black} further in Section \ref{sec:discussion}.

@@

The physical parameter training sample consists of 87 sources with reference labels distributed over $2859 <T_{\rm eff} < 4131$K, and $-0.48 < \feh < 0.49$\,dex, 41 sources of which are drawn from \citet{Mann:2015} (hereafter M15), and 46 of which are part of a previously unpublished extension sample to M15, analyzed using similar data and identical techniques to M15. The major difference between the extension sample is that the sample had lower-quality or no parallaxes (prior to Gaia data) and hence were omitted from the M15 study and were less vetted for binarity than the M15 sample \color{red}{\bf (however, all sources in the training sample were visually inspected by color-magnitude position for binarity before addition).}\color{black}

@@

Another mode of analysis we can \color{red}{\bf utilize }\color{black} with \thecannon\ is how the derivative of the model changes with respect to given training parameters, which makes our model interpretable for discovering or verifying atomic or molecular lines with strong dependence on different physical parameters. The top two panels of Figures \ref{fig:demo_teff} and \ref{fig:demo_feh} show two example spectra and model fits for two different temperatures (Figure \ref{fig:demo_teff}), and two different metallicities (Figure \ref{fig:demo_feh}), with atomic and molecular features identified by the abundance analysis of \citet{Souto:2017}. The bottom panels of Figures \ref{fig:demo_teff} and \ref{fig:demo_feh} show the derivative of flux with respect to temperature and metallicity at each pixel, taken at the median training values. In order to evaluate which spectral features show statistically significant change with respect to input label, we compute \color{red}{\bf the }\color{black} error of the derivative at each pixel using a jackknife statistic (with a 1-$\sigma$ level overplotted in red):
\begin{equation}
	\sigma_{\theta,m}^2 = \frac{N-1}{N} \sum^N_{n=1} (\theta_{/n}^m - \theta^m)^2 
\end{equation}
where $\sigma_{\theta,m}$ is the error at pixel \emph{m}, \emph{N} is the total number of stars in the sample indexed by \emph{n}, $\theta$ is the coeffient vector trained on all \emph{N} sources, and $\theta_{/n}$ is the coefficient vector trained on \emph{N-1} sources excluding star \emph{n}. A summary of identified lines with derivative values greater than $2\sigma_{\rm jackkife}$ is given in Table \ref{table:teff_derivative} and Table \ref{table:feh_derivative}.

@@

The spectra contain roughly 8000 pixels,
so we might expect the $\chi^2$ values to be close to 8000 in magnitude,
\color{red}{\bf however, they }\color{black} are much higher.
This discrepancy flows from the fact that, while the spectral model is good at the percent
level, the signal-to-noise of the typical spectra are more than 100.
That is, the $\chi^2$ values do show that the model is not good in the frequentist
sense; it is only good at the percent level.

@@

We trained \thecannon\ on 51 M dwarfs ranging \color{red}{\bf between }\color{black} M0$-$M9 with a one-dimensional spectral type label, and obtained a precision of $\pm$0.9 spectral types, also similar to the original training label uncertainty of $\pm$1 spectral type. We note however, that the training sample is distributed heavily towards earlier type sources, with a median spectral type of 3 and only one M8 and one M9 source. As seen in Figure \ref{fig:west_validation}, the model performs poorly at reproducing spectral types $>$M8, which confirms that \thecannon\ does not extrapolate well to labels outside of the training sample space. Because of this skew for late-type sources, we report our spectral type model to be precise to $\pm$0.7 spectral types for the range M0-M6. Repeating the analysis of Section \ref{subsec:mann_results}, Figure \ref{fig:west_validation} shows LOOCV test for the labels reported in Table \ref{table:west_results}, and Figure \ref{fig:west_derivative} shows the derivative of model flux with varying spectral type.

@@

\begin{enumerate}
\item \textbf{Quality of fit cut:} We apply a \cannon\ model $\chi^2$ cut of less than 100,000, chosen to remove badly fit sources (such as fast rotators), but include $\chi^2$ values close \color{red}{\bf to }\color{black} the distribution of training sample (Figure \ref{fig:chi_dist}).

\item \textbf{Color-magnitude cuts:} Using \gaia\ and 2MASS photometry we apply the \color{red}{\bf additional }\color{black} color-magnitude selections shown in Figure \ref{fig:cmd_selection} to remove sources above the main sequence (which are likely pre-main sequence, reddened K dwarfs and\color{red}{\bf /or }\color{black} multiples), and subdwarfs below the main sequence.

@@

\begin{figure}
\begin{center}
	\includegraphics[width=\linewidth]{wide_binary_met.png}
	\caption{Metallicities of \cannon\ M dwarfs in binaries (x-axis) compared to the metallicities of their companions (y-axis). Label colors show the source and spectral type of the companion metallicities. Metallicity values for each binary pair is compiled in Table \ref{table:binaries}. The overall scatter between the 15 metallicity pairs is 0.08\,dex.} 
	\label{fig:wide_binary_met}
\end{center}
\end{figure}

@@

As a check of our test sample metallicity reliability, we cross-matched our M dwarf final sample with the \citet{Elbadry:2018b} catalog of $>$50,000 high-confidence, widely separated binaries identified by \drtwo. 
In total we found 216 of the APOGEE M dwarfs to have binary pairs (46 FGK+M, 155 M+M, and 15 WD+M). Out of the 155 M+M pairs, 8 pairs contained both pairs in APOGEE. Cross-matching the list of FGK+M dwarf companions with several catalogs/surveys with measured stellar metallicities, we found an additional 7 sources with FGK metallicities from LAMOST \citep{Zhao:2012} and APOGEE (ASPCAP). The metallicity measurements for the 15 M dwarf binaries and their companions are given in Table \ref{table:binaries} \color{red}{\bf and shown in Figure \ref{fig:wide_binary_met} }\color{black}, and the overall scatter is $0.08\,$dex--an improvement over the scatter of ASPCAP metallicities which is $0.15\,$dex for these 15 sources. Internal consistency of the two models (the scatter of the 8 M+M pairs both in APOGEE) are 0.06\,dex and 0.12\,dex for \cannon\ and ASPCAP respectively.

@@

While \thecannon\ achieves excellent precision at predicting labels
and reproducing spectral features, the
accuracy of labels it produces is limited by
the accuracy, relative precision, size, dynamic range and
representation of the training sample.
That is, being a supervised method, \thecannon\ is never any better in a mean (bias) sense than the input training data, although it can be better in a precision or variance sense.
The catalog we have produced is a label transfer from parameters provided
in our input data (M15) and it implicitly adopts all the biases and
issues from that input data.
It is also limited to the stellar-parameter domain of that input catalog.
That said, this work provides an external validation of the M15 stellar parameters.

@@

While a detailed analysis of atmospheric model limitations is beyond the scope of this paper, 
our results provide an avenue to compare the FGK star metallicity scale 
to the less well-understood M dwarf metallicity scale.
These results find that atmospheric metallicities are systematically metal-poor biased
compared to \cannon-based metallicities trained on sources with metallicities calibrated
to those of FGK companions. 
At the high temperature end ($\teff > 3600$\,K), the ASPCAP metallicity bias is $-0.12-0.15\,$dex with a scatter of $0.05-0.06\,$dex relative to \cannon\ metallicities, and increases to a bias of $-0.3\,$dex and scatter of $0.13\,$dex at the lower temperature end ($\teff < 3200$\,K) (Figure \ref{fig:met_comp}).
We suspect that this metal-poor bias, while not explored to a great extent in this work, is due 
to the line lists of the models: 
an effect in which the optimizer of the pipeline may be lowering the continuum level and metallicity of the fit to compensate for the missing lines or opacities.
\color{red}{\bf We also note that this analysis was completed using data from Data Release 14 of APOGEE, 
which did include molecular lines from FeH in the pipeline at the time, which become numerous and strong for $\teff \lesssim 3600$\,K. }\color{black}
Further analysis would need to be done to \color{red}{\bf quantify the metallicity improvement for M dwarfs in later data releases of APOGEE now available (in which FeH has been included in analysis)}\color{black}, and determine if the metallicity bias
is found in other model grids (besides the ATLAS/MARCS models used by the ASPCAP pipeline), 
and if the effect is present at other wavelengths.

@@

%%==========================================================================================
\acknowledgements
We would like to acknowledge Hans Walter Rix (MPIA), Derek Homeier (Universit{\"a}t Heidelberg), Wolfgang Bradner (MPIA), Anna-Christina Eilers (MPIA), Melissa Ness (Columbia), Kevin Covey (WWU), Diogo Souto (Observatário Nacional/MCTI), Keivan Stassun (Vanderbilt), \color{red}{\bf Katia Cunha (NOAO), Anthony Brown (Leiden), Aida Behmard (Caltech) and Christopher Theissen (UCSD) }\color{black} for constructive discussions in the process of this project, as well as Bertrand Goldman (MPIA) and those who have supported the internship program at the Max Planck Institute f{\"u}r Astronomie for providing JB with funding and hospitality.
This project was partially supported by
the US National Aeronautics and Space Administration (NASA grant NNX12AI50G),
the US National Science Foundation (NSF grant AST-1517237),
and the Moore--Sloan Data Science Environment at NYU.
A.J.B. acknowledges funding support from the National Science Foundation under award No. AST-1517177.
This work is supported by the SDSS Faculty and Student Team (FAST) initiative.

@@

\begin{figure*}
	\begin{center}
	\includegraphics[width=16cm]{demo_derivatives_teff3.png}
	\end{center}
	\caption{\textit{Top two panels of each plot:} \color{red}{\bf APOGEE spectra (black), overlaid by the }\color{black} Mann-trained \color{red}{\bf \cannon\ }\color{black} model for \color{red}{\bf two sources of }\color{black} varying temperatures, and similar metallicities. \textit{Third panel of each plot:} Derivative of \thecannon\ model with respect to temperature, taken at the median training temperature, T$_{\rm eff}=3463$K; error estimate computed using a jackknife statistic at each pixel is marked in red, making it possible to distinguish which features vary significantly with change in spectral type, and which are likely due to noise.}
	\label{fig:demo_teff}
\end{figure*}

@@

\begin{figure*}
	\begin{center}
	\includegraphics[width=16cm]{demo_derivatives_feh3.png}
	\end{center}
	\caption{\textit{Top two panels of each plot:} \color{red}{\bf APOGEE spectra (black), overlaid by the }\color{black} Mann-trained \color{red}{\bf \cannon\ }\color{black} model for \color{red}{\bf two sources of }\color{black} varying metallicities and similar temperatures. \textit{Third panel:} Derivative of \thecannon\ model with respect to metallicity, taken at the median training metallicity, $\feh=-0.03$\,dex; jackknife computed error at each pixel is shown in red.} \label{fig:demo_feh}
\end{figure*}
